---
title: "Elo Merchant Category Recommendation Preprocessing"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Read in data.
```{r}
rm(list = ls())
library(tidyverse)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(lubridate)
library(imputeMissings)

train = read_delim("train.csv", delim = ',')
test = read_delim("test.csv", delim = ',')
merchants = read_delim("merchants.csv", delim = ',')
new_trans = read_delim("new_merchant_transactions.csv", delim = ',')
hist_trans = read_delim("historical_transactions.csv", delim = ',')
```

## Factorize and Create New Columns

Historical Transactions
```{r}
hist_trans$authorized_flag = ifelse(hist_trans$authorized_flag == "Y", 1, 0)
hist_trans$year = year(hist_trans$purchase_date)
hist_trans$month = month(hist_trans$purchase_date)
hist_trans$day = day(hist_trans$purchase_date)
hist_trans$hour = hour(hist_trans$purchase_date)
hist_trans$minute = minute(hist_trans$purchase_date)
hist_trans$second = second(hist_trans$purchase_date)
```

New Transactions
```{r}
new_trans$authorized_flag = ifelse(new_trans$authorized_flag == "Y", 1, 0)
new_trans$year = year(new_trans$purchase_date)
new_trans$month = month(new_trans$purchase_date)
new_trans$day = day(new_trans$purchase_date)
new_trans$hour = hour(new_trans$purchase_date)
new_trans$minute = minute(new_trans$purchase_date)
new_trans$second = second(new_trans$purchase_date)
```


## Remove Values that are Improperly Recorded (such as -1, 999, etc.) and Outliers in Purchase Amount
```{r}
summary(hist_trans)
```

Things to remove: high number of installments (> 12), high (normalized) purchase amount (> 1), id values of -1 (city_id, merchant_category_id, state_id, subsector_id)

```{r}
hist_trans[hist_trans$installments > 12 | hist_trans$installments < 0, "installments"] = NA
hist_trans[hist_trans$purchase_amount > 1, "purchase_amount"] = NA
hist_trans[hist_trans$city_id == -1, "city_id"] = NA
hist_trans[hist_trans$merchant_category_id == -1, "merchant_category_id"] = NA
hist_trans[hist_trans$state_id == -1, "state_id"] = NA
hist_trans[hist_trans$subsector_id == -1, "subsector_id"] = NA
```

```{r}
summary(new_trans)
```

```{r}
new_trans[new_trans$installments > 12 | new_trans$installments < 0, "installments"] = NA
new_trans[new_trans$purchase_amount > 1, "purchase_amount"] = NA
new_trans[new_trans$city_id == -1, "city_id"] = NA
new_trans[new_trans$merchant_category_id == -1, "merchant_category_id"] = NA
new_trans[new_trans$state_id == -1, "state_id"] = NA
new_trans[new_trans$subsector_id == -1, "subsector_id"] = NA
```

## Missing Value Imputation 

Historical transactions
```{r}
sapply(hist_trans, function(x) any(is.na(x)))
```

```{r}
hist_trans = impute(hist_trans)
```

New Transactions
```{r}
sapply(new_trans, function(x) any(is.na(x)))
```

```{r}
new_trans = impute(new_trans)
```

## Aggregation

Create mode function
```{r}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```

Things to aggregate by group of functions:
- card_id: tally
- authorized_flag: mean, sd, mode, median, min, max, n_distinct
- installments: mean, sd, mode, median, min, max, n_distinct
- month_lag: mean, sd, mode, median, min, max, n_distinct,
- purchase_amount: mean, sd, mode, median, min, max, n_distinct
- city_id: n_distinct, mode
- merchant_category_id: n_distinct, mode
- merchant_id: n_distinct, mode
- state_id: n_distinct, mode
- subsector_id: n_distinct, mode
- category_1: n_distinct, mode
- category_2: n_distinct, mode
- category_3: n_distinct, mode
- year: n_distinct, mode, min, max
- month: n_distinct, mode, min, max
- day: n_distinct, mode, min, max
- hour: n_distinct, mode, min, max
- minute: n_distinct, mode, min, max
- second: n_distinct, mode, min, max
- purchase_date: latest, earliest, latest - earliest

Store collection of columns
```{r}
var_1 = c("authorized_flag", "installments", "month_lag", "purchase_amount")
var_2 = c("city_id", "merchant_id", "merchant_category_id", "state_id", "subsector_id")
var_3 = c("year", "month", "day", "hour", "minute", "second")
var_4 = c("category_1", "category_2", "category_3")
```

For new transactions:
```{r}
nt_0 = new_trans %>% group_by(card_id) %>% 
  count(card_id) %>% ungroup()
nt_1 = new_trans %>% select(card_id, var_1) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_1, 
               .funs = c("mean", "sd", "Mode",
                         "median", "min", "max", "n_distinct")) %>% ungroup()
nt_2 = new_trans %>% select(card_id, var_2) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_2, 
               .funs = c("n_distinct", "Mode")) %>% ungroup() 
nt_3 = new_trans %>% select(card_id, var_3) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_3, 
               .funs = c("n_distinct", "Mode", 
                         "min", "max")) %>% ungroup() 
nt_4 = new_trans %>% select(card_id, var_4) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_4, 
               .funs = c("n_distinct", "Mode")) %>% ungroup() 
nt_5 = new_trans %>% select(card_id, purchase_date) %>% group_by(card_id) %>% 
  mutate("purchase_date_max" = max(purchase_date), 
         "purchase_date_min" =  min(purchase_date), 
         "purchase_date_range" = as.numeric(difftime(max(purchase_date), 
                                                     min(purchase_date), 
                                                     units = "d"))) %>% 
  filter(row_number() == 1)
```

Put columns together and keep distinct rows 
```{r}
new_trans_overall = left_join(nt_0, nt_1, on = card_id) %>% 
  left_join(., nt_2, on = card_id) %>% 
  left_join(., nt_3, on = card_id) %>% 
  left_join(., nt_4, on = card_id) %>% 
  left_join(., nt_5, on = card_id) %>% 
  distinct(.,card_id, .keep_all= TRUE)
```

```{r}
which(sapply(new_trans_overall, function(x) any(is.na(x))))
```

Replace null standard deviations with 0
```{r}
new_trans_overall$authorized_flag_sd = new_trans_overall$authorized_flag_sd %>% replace_na(0)
new_trans_overall$installments_sd = new_trans_overall$installments_sd %>% replace_na(0)
new_trans_overall$month_lag_sd = new_trans_overall$month_lag_sd %>% replace_na(0)
new_trans_overall$purchase_amount_sd = new_trans_overall$purchase_amount_sd %>% replace_na(0)
```

Save aggregated new transactions file
```{r}
write.csv(new_trans_overall, file = "new_transactions_explained.csv")
rm(nt_0, nt_1, nt_2, nt_3, nt_4, nt_5, new_trans, new_trans_overall)
```

For nistorical transactions: 
```{r}
ht_0 = hist_trans %>% group_by(card_id) %>% 
  count(card_id) %>% ungroup() 
ht_1 = hist_trans %>% select(card_id, var_1) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_1, 
               .funs = c("mean", "sd", "Mode", 
                         "median", "min", "max", "n_distinct")) %>% ungroup() 
ht_2 = hist_trans %>% select(card_id, var_2) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_2, 
               .funs = c("n_distinct", "Mode")) %>% ungroup() 
ht_3 = hist_trans %>% select(card_id, var_3) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_3, 
               .funs = c("n_distinct", "Mode", 
                         "min", "max")) %>% ungroup() 
ht_4 = hist_trans %>% select(card_id, var_4) %>% group_by(card_id) %>% 
  summarize_at(.vars = var_4, 
               .funs = c("n_distinct", "Mode")) %>% ungroup() 
ht_5 = hist_trans %>% select(card_id, purchase_date) %>% group_by(card_id) %>% 
  mutate("purchase_date_max" = max(purchase_date), 
         "purchase_date_min" =  min(purchase_date), 
         "purchase_date_range" = as.numeric(difftime(max(purchase_date), 
                                                     min(purchase_date), 
                                                     units = "d"))) %>% 
  filter(row_number() == 1)
```

Put columns together and keep distinct rows
```{r}
hist_trans_overall = left_join(ht_0, ht_1, on = card_id) %>% 
  left_join(., ht_2, on = card_id) %>% 
  left_join(., ht_3, on = card_id) %>% 
  left_join(., ht_4, on = card_id) %>% 
  left_join(., ht_5, on = card_id) %>% 
  distinct(.,card_id, .keep_all= TRUE)
```

```{r}
which(sapply(hist_trans_overall, function(x) any(is.na(x))))
```

Save aggregated historical transactions file
```{r}
write.csv(hist_trans_overall, file = "historical_transactions_explained.csv")
rm(ht_0, ht_1, ht_2, ht_3, ht_4, ht_5, hist_trans, hist_trans_overall)
```
